### 2016-06-28 Projects/DelCo_2014_2015/Reports/Hose_Stream_Tests/TeX_Files/DelCo_Hose_Stream_Tests.tex
Dan's final edits to hose stream report; updated HRRs & construction description in burner report

Fire suppression tactics using hose streams can affect ventilation in a structure and may impact the movement of smoke and heat through a structure. Seven series of experiments were conducted to study the impact of different hose stream and nozzle movement pattern combinations on air movement within residential scale structures. The experiments studied four different hose streams: a straight stream, narrow fog stream, and a wide fog stream from a combination nozzle and a solid stream from a smooth bore nozzle. The streams were applied in a static, or fixed, position; by moving the hoseline left-to-right across the room in a sweeping motion; and by rotating the hoseline in both the clockwise and counterclockwise directions. Gas velocity was measured at different locations in the structure during the experiments. The wide fog stream caused the most air movement out of any of the tested streams, reaching a maximum of 2.6~m/s, followed by the narrow fog stream, and then the straight stream and solid stream from the smooth bore nozzle, which both caused approximately the same amount of air movement through the structure. It was discovered that the straight stream only caused significant air movement through the structure when it was applied in a moving pattern. Furthermore, there were no statistically significant differences in the average measured air velocity between the clockwise and counterclockwise nozzle movement patterns. It was determined that the type of hose stream and manner in which it is applied dictates the extent to which a stream impacts the ventilation of a structure. To better quantify the impact of such air flows on the fire environment, additional fire experiments need to be conducted with structure fires in a controlled environment.

@@

\begin{enumerate}
	\item Determine which hose stream patterns and nozzle movement patterns cause significant air movement in a structure 
	\item Determine the amount of air movement induced by different hose streams and nozzle movement patterns

\end{enumerate}

@@

Looking at the table and figures, a number of trends can be observed. For example, there was significant air movement---an average gas velocity greater than 0.5~m/s (1.1~mph)---through the interior doorway while the flow path was fully established for every hose stream and target location combination except the straight stream in Test~1. Looking at Table~\ref{table:all_mon_vel_avgs}, the narrow fog and wide fog streams caused higher velocities through the interior doorway when the flow path was fully established compared to the straight stream. This trend can also be seen graphically in Fig.~\ref{fig:Test_2_BDP_A10_Avg_All} and Figs.~\ref{fig:Test_1_BDP_A6_Avg_All}~and~\ref{fig:Test_3_BDP_A10_Avg_All} in Appendix~\ref{chap:monitor_plots}. Furthermore, the wide fog stream generally produced equal or higher air velocities compared to the narrow fog stream. Finally, the Test~4 results listed in Table~\ref{table:all_mon_vel_avgs} and plotted in Fig.~\ref{fig:Test_4_BDP_A10_Avg_All} suggest that the straight stream from a combination nozzle and the solid stream from a smooth bore nozzle equipped with a 1~in tip produced approximately the same amount of air movement through the interior doorway.
\FloatBarrier

@@

With the exception of the narrow fog stream in the fixed position during Test~6, every hose stream and nozzle movement pattern combination using a narrow fog or wide fog stream during the handline experiments produced significant air movement---an average gas velocity greater than 0.5~m/s (1.1~mph)---through the interior doorway while the flow path was fully established. For the straight stream, however, significant air movement through the interior doorway only occurred when the stream was applied in a pattern that involved moving the handline. Similar to the monitor experiments, the average air velocity through the interior doorway was always greater in magnitude for the nozzle movement patterns using the narrow and wide fog streams than the same patterns using the straight stream. Another important result seen in Tables~\ref{table:east_hand_A6_avgs} and \ref{table:west_hand_A10_avgs} is that there are no statistically significant differences in the amount of air flow through the interior doorway when rotating the hoseline in the clockwise direction compared to rotating it in the counterclockwise direction for any of three hose streams tested. The similarities in air flow produced by the clockwise and counterclockwise nozzle movement patterns can also be seen graphically in Fig.~\ref{fig:Test_6_BDP_A10_Avg_All} and Figs.~\ref{fig:Test_5_BDP_A6_Avg_All} and \ref{fig:Test_7_BDP_A10_Avg_All} in Appendix~\ref{chap:handline_plots}.

@@

In addition to the monitor test series, three different experimental series were conducted using a handline to study the impact of different stream and nozzle movement pattern combinations on air movement within a structure. Similar to the monitor experiments, the average air velocity through the interior doorway of a structure with a fully established flow path was greater for the wide fog and narrow fog streams than the straight stream. In fact, it was discovered that a straight stream only caused significant air movement through the interior doorway when it was applied in a moving pattern. Additionally, throughout all handline test series, there were no statistically significant differences in the average air velocity measured through the interior doorway between the clockwise and counterclockwise nozzle movement patterns.

Overall, the water flow experiments discussed in this report reveal that hose streams can affect air movement within a structure. The type of hose stream used and manner in which it is applied dictates the extent to which a stream impacts the ventilation of a structure. These experiments have shown there is a potential for hose streams to affect air flows within a structure during fire scenarios. To better quantify the impact of air flows from hose streams on the fire environment, additional fire experiments need to be conducted with structure fires in a controlled environment.

>>>>

Fire suppression tactics using hose streams can affect ventilation in a structure and may impact the movement of smoke and heat through a structure. Seven series of experiments were conducted to study the impact of different hose stream and nozzle movement pattern combinations on air movement within residential scale structures. The experiments studied four different hose streams: a straight stream, narrow fog stream, and a wide fog stream from a combination nozzle and a solid stream from a smooth bore nozzle. The streams were applied from a static, or fixed, position; by moving the hoseline left-to-right across the room in a sweeping motion; and by rotating the hoseline in both the clockwise and counterclockwise directions. Gas velocity was measured at different locations in the structure during the experiments. The wide fog stream caused the most air movement out of any of the tested streams, reaching a maximum velocity of 2.6~m/s (5.8~mph) and maximum air flow rate of approximately 4.2~m$^3$/s (9000~cfm), followed by the narrow fog stream, and then the straight stream and solid stream from the smooth bore nozzle, which both caused approximately the same amount of air movement through the structure. It was discovered that the straight stream only caused air movement through the structure when it was applied in a moving pattern. Furthermore, there were no statistically significant differences in the average measured air velocity between the clockwise and counterclockwise nozzle movement patterns. It was determined that the type of hose stream and manner in which it is applied dictates the extent to which a stream impacts the ventilation of a structure. To better quantify the impact of such air flows on the fire environment, additional experiments need to be conducted with structure fires in a controlled environment.

@@

\begin{enumerate}
	\item Determine the amount of air movement induced by different hose streams and nozzle movement patterns.
	\item Examine Royer's nozzle movement theory on the direction of nozzle rotation.
\end{enumerate}

@@

Looking at the table and figures, a number of trends can be observed. For example, there was air movement---an average gas velocity greater than 0.5~m/s (1.1~mph)---through the interior doorway while the flow path was fully established for every hose stream and target location combination except the straight stream in Test~1. Looking at Table~\ref{table:all_mon_vel_avgs}, the narrow fog and wide fog streams caused higher velocities through the interior doorway when the flow path was fully established compared to the straight stream. This trend can also be seen graphically in Fig.~\ref{fig:Test_2_BDP_A10_Avg_All} and Figs.~\ref{fig:Test_1_BDP_A6_Avg_All}~and~\ref{fig:Test_3_BDP_A10_Avg_All} in Appendix~\ref{chap:monitor_plots}. Furthermore, the wide fog stream generally produced equal or higher air velocities compared to the narrow fog stream. Finally, the Test~4 results listed in Table~\ref{table:all_mon_vel_avgs} and plotted in Fig.~\ref{fig:Test_4_BDP_A10_Avg_All} suggest that the straight stream from a combination nozzle and the solid stream from a smooth bore nozzle equipped with a 1~in tip produced approximately the same amount of air movement through the interior doorway.
\FloatBarrier

@@

With the exception of the narrow fog stream in the fixed position during Test~6, every hose stream and nozzle movement pattern combination using a narrow fog or wide fog stream during the handline experiments produced air movement---an average gas velocity greater than 0.5~m/s (1.1~mph)---through the interior doorway while the flow path was fully established. For the straight stream, however, air movement through the interior doorway only occurred when the stream was applied in a pattern that involved moving the handline. Similar to the monitor experiments, the average air velocity through the interior doorway was always greater in magnitude for the nozzle movement patterns using the narrow and wide fog streams than the same patterns using the straight stream. Another important result seen in Tables~\ref{table:east_hand_A6_avgs} and \ref{table:west_hand_A10_avgs} is that there are no statistically significant differences in the amount of air flow through the interior doorway when rotating the hoseline in the clockwise direction compared to rotating it in the counterclockwise direction for any of three hose streams tested. The similarities in air flow produced by the clockwise and counterclockwise nozzle movement patterns can also be seen graphically in Fig.~\ref{fig:Test_6_BDP_A10_Avg_All} and Figs.~\ref{fig:Test_5_BDP_A6_Avg_All} and \ref{fig:Test_7_BDP_A10_Avg_All} in Appendix~\ref{chap:handline_plots}.

@@

In addition to the monitor test series, three different experimental series were conducted using a handline to study the impact of different stream and nozzle movement pattern combinations on air movement within a structure. Similar to the monitor experiments, the average air velocity through the interior doorway of a structure with a fully established flow path was greater for the wide fog and narrow fog streams than the straight stream. It was discovered that a straight stream increased air movement through the interior doorway when it was applied in a moving pattern. With regard to Royer's theory, throughout all handline test series, there were no statistically significant differences in the average air velocity measured through the interior doorway between the clockwise and counterclockwise nozzle movement patterns. However, these experiments only addressed cold flow.

Overall, the water flow experiments discussed in this report reveal that hose streams can affect air movement within a structure. The type of hose stream used and manner in which it is applied dictates the extent to which a stream impacts the ventilation of a structure. These experiments have shown there is a potential for hose streams to affect air flows within a structure during fire scenarios. To better quantify the impact of air flows from hose streams on the fire environment, additional experiments need to be conducted with structure fires in a controlled environment.

### 2016-06-28 Projects/DelCo_2014_2015/Reports/Propane_Gas_Fire_Experiments/TeX_Files/Propane_Gas_Fire_Experiments.tex
Dan's final edits to hose stream report; updated HRRs & construction description in burner report

The interior walls of the first floor of each structure were framed with steel studs set to 400~mm (16~in) centers and track and were lined with 13~mm (0.5~in) thick cement board. The walls were composed of 16~mm (0.63~in) Type X gypsum board. Additionally, the ceiling was composed of two layers of 13~mm (0.5~in) thick cement board.
\FloatBarrier

@@

\subsubsection{Second Floor of West Structure}
The second floor of the West Structure was built on the wood ceiling support described above and was connected to the first floor by a stairwell. The second story walls were of wood-frame with 51~mm (2~in) by 102~mm (4~in) studs set to 400~mm (16~in) centers. The interior walls were protected by 16~mm (0.63~in) fire rated gypsum board, 16~mm (0.63~in) durarock board, and a second layer of 16~mm (0.63~in) fire rated gypsum board. The exterior walls were protected with 11~mm (0.44~in) oriented strand board and 8~mm (0.31~in) fiber cement lap siding.

@@

\begin{table}[!ht]
\caption{Heat Release Rates for Tests~2--4}
\begin{tabular}{lccccc}
 \toprule
 					& 	\multicolumn{5}{c}{\textbf{Heat Release Rate (kW)}} \\
\textbf{Test} 		&	\begin{tabular}{c} Corner \\ Burner On \end{tabular} 
					&	\begin{tabular}{c} Middle \\ Burner On \end{tabular}
					&	\begin{tabular}{c} Center \\ Burner On \end{tabular}
					& 	\begin{tabular}{c} Center \\ Burner Off \end{tabular}
					&   \begin{tabular}{c} Middle \\ Burner Off \end{tabular} \\
 \midrule
Test 2 				& 340 & 650	& 900 & 630 & 330 \\
Test 3 				& 330 & 650	& 910 & 630 & 330 \\
Test 4				& 350 & 650	& 950 & 650 & 340 \\
\bottomrule
\end{tabular}
\label{table:HRR_Tests_2-4}
\end{table}

@@

\begin{table}[!ht]
\caption{Heat Release Rates for Tests~5 and 6}
\begin{tabular}{lc}
 \toprule
 & \textbf{Heat Release Rate (kW)} \\
\textbf{Test} & \textbf{All Burners On} \\
\midrule
Test 5 - Seq. 1		& 730 \\
Test 5 - Seq. 2		& 730 \\
Test 5 - Seq. 3		& 720 \\
Test 6 - Seq. 1		& 720 \\
Test 6 - Seq. 2		& 720 \\
Test 6 - Seq. 3		& 710 \\
\bottomrule
\end{tabular}
\label{table:HRR_Tests_5-6}
\end{table}

@@

\begin{table}[!ht]
\caption{Heat Release Rates for West Structure Tests}
\begin{tabular}{lc}
 \toprule
					& 	\textbf{Heat Release Rate (kW)}	\\
\textbf{Test}		& All burners on \\
 \midrule
Test 22				&     	760 	  \\
Test 23				&     	790 	  \\
Test 24				& 	    790 	  \\
Test 25				&     	790 	  \\
\bottomrule
\end{tabular}
\label{table:HRR_West}
\end{table}

>>>>

The interior walls of the first floor of each structure were framed with steel studs set to 400~mm (16~in) centers and track. Two layers of 16~mm (0.63~in) Type X gypsum board lined the steel studs, and a layer of 13~mm (0.5~in) thick cement board covered the gypsum board. Additionally, the ceiling was composed of two layers of 13~mm (0.5~in) thick cement board.
\FloatBarrier

@@

\subsubsection{Second Floor of West Structure}
The second floor of the West Structure was built on the wood ceiling support described above and was connected to the first floor by a stairwell. The second story walls were of wood-frame with 51~mm (2~in) by 102~mm (4~in) studs set to 400~mm (16~in) centers. Two layers of 16~mm (0.63~in) Type X gypsum board lined the wood studs, and a layer of 13~mm (0.5~in) thick cement board covered the gypsum board. The exterior walls were protected with 11~mm (0.44~in) oriented strand board and 8~mm (0.31~in) fiber cement lap siding.

@@

\begin{table}[!ht]
\caption{Heat Release Rates for Tests~2--4}
\begin{tabular}{lccccc}
 \toprule
 					& 	\multicolumn{5}{c}{\textbf{Heat Release Rate (kW)}} \\
\textbf{Test} 		&	\begin{tabular}{c} Corner \\ Burner On \end{tabular} 
					&	\begin{tabular}{c} Middle \\ Burner On \end{tabular}
					&	\begin{tabular}{c} Center \\ Burner On \end{tabular}
					& 	\begin{tabular}{c} Center \\ Burner Off \end{tabular}
					&   \begin{tabular}{c} Middle \\ Burner Off \end{tabular} \\
 \midrule
Test 2 				& 560 & 1080 & 1470 & 1070 & 550 \\
Test 3 				& 530 & 1040 & 1470 & 1050 & 520 \\
Test 4				& 580 & 1140 & 1580 & 1130 & 580 \\
\bottomrule
\end{tabular}
\label{table:HRR_Tests_2-4}
\end{table}

@@

\begin{table}[!ht]
\caption{Heat Release Rates for Tests~5 and 6}
\begin{tabular}{lc}
 \toprule
 & \textbf{Heat Release Rate (kW)} \\
\textbf{Test} & \textbf{All Burners On} \\
\midrule
Test 5 - Seq. 1		& 1190 \\
Test 5 - Seq. 2		& 1190 \\
Test 5 - Seq. 3		& 1190 \\
Test 6 - Seq. 1		& 1190 \\
Test 6 - Seq. 2		& 1190 \\
Test 6 - Seq. 3		& 1180 \\
\bottomrule
\end{tabular}
\label{table:HRR_Tests_5-6}
\end{table}

@@

\begin{table}[!ht]
\caption{Heat Release Rates for West Structure Tests}
\begin{tabular}{lc}
 \toprule
					& 	\textbf{Heat Release Rate (kW)}	\\
\textbf{Test}		& All burners on \\
 \midrule
Test 22				&     	1240 	  \\
Test 23				&     	1290 	  \\
Test 24				& 	    1270 	  \\
Test 25				&     	1270 	  \\
\bottomrule
\end{tabular}
\label{table:HRR_West}
\end{table}
